---
title: "Risk Premia Strategy Notebook"
output: html_notebook
---

Pull data from bigquery

```{r}
if (!require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse, bigrquery, slider, lubridate, tidyquant, here)

prices <- query_exec(
  "SELECT *  FROM `rw-algotrader.master_assetclass.assetclass_price` WHERE ticker in ('VTI','GLD','TLT');",
  project = 'rw-algotrader',
  use_legacy_sql = FALSE
)

tb <- bq_project_query(
  'rw-algotrader',
  "SELECT *  FROM `rw-algotrader.master_assetclass.assetclass_price` WHERE ticker in ('VTI','GLD','TLT');"
)
bq_table_fields(tb)
p <- bq_table_download(tb)


load(here::here("data", "assetclass_prices.RData"))

prices <- prices %>% arrange(date) %>% as_tibble()
```

Plot cumulative returns to verify things...

```{r}
prices %>%
  arrange(date) %>%
  group_by(ticker) %>%
  mutate(returns = closeadjusted / dplyr::lag(closeadjusted) - 1) %>%
  na.omit() %>%
  mutate(cumreturns = cumprod(1+returns)) %>%
   ggplot(aes(x=date, y = cumreturns, color = ticker)) + geom_line()
```

# Backtest setup

Calculate monthly prices, and start and end dates...

```{r}
monthends <- prices %>% 
  mutate(year = year(date),
         month = month(date)) %>%
  group_by(year, month) %>%
  summarise(date = max(date))

monthlyprices <- prices %>%
  inner_join(monthends, by = 'date') %>%
  select(ticker, date, close = closeadjusted)

startDate <- monthlyprices %>% summarise(min(date)) %>% pull()
endDate <- monthlyprices %>% summarise(max(date)) %>% pull()
initDate <- startDate - 1 # InitDate is day before startdate
```


# Tab 1: Equal Initial Weight Buy and Hold

Here are the inputs that should be on this tab:

```{r}
initEq <- 10000 # Allow $1000 to 1 million
perShareComm <- 0.005 # Allow 0.001 to 0.02 
minCommPerOrder <- 1 # Allow 0 to 5
```

This is basically assuming that rebalFrequency = capFrequency = 0.

```{r}
shares <- monthlyprices %>%
  filter(date == startDate) %>%
  mutate(shares = trunc((initEq / 3) / close)) %>%
  select(ticker, shares)
  
ew_norebal <- monthlyprices %>%
  inner_join(shares, by = 'ticker') %>%
  mutate(exposure = shares * close) %>%
  group_by(ticker) %>%
  mutate(trades = shares - lag(shares)) %>%
  # Initial trade to setup position comes through as NA
  mutate(trades = case_when(is.na(trades) ~ shares, TRUE ~ trades)) %>%
  mutate(tradevalue = trades * close) %>%
  mutate(commission = case_when(abs(trades) * perShareComm > minCommPerOrder ~ abs(trades) * perShareComm, trades == 0 ~ 0, TRUE ~ minCommPerOrder))

# Calculate cash balance
initcashbal <- ew_norebal %>% 
  ungroup() %>%
  filter(date == startDate) %>%
  summarise(cash = initEq - sum(exposure) - sum(commission)) %>%
  pull()


cash <- ew_norebal %>%
  ungroup() %>%
  filter(ticker == 'VTI') %>% # just doing this to get the dates
  mutate(ticker = 'Cash', 
         date, 
         close = 0, 
         shares = 0, 
         exposure = initcashbal, 
         trades = 0, 
         tradevalue = case_when(date == startDate ~ initcashbal - initEq, TRUE ~ 0), commission = 0)

# Bind cash balances and sort by date again
ew_norebal <- ew_norebal %>%
  bind_rows(cash) %>%
  arrange(date)

# Stacked area chart
ew_norebal %>% 
  ggplot(aes(x=date, y=exposure, fill=ticker)) +
    geom_area() + 
    theme_set(theme_bw()) + 
    ggtitle('3 ETF USD Risk Premia - Equal Weight, No Rebalancing')

# Trades chart
ew_norebal %>%
  ggplot(aes(x=date, y=tradevalue, fill=ticker)) +
    geom_bar(stat = 'identity', position = position_dodge()) + 
    theme_set(theme_bw()) + 
    ggtitle('3 ETF USD Risk Premia - Trades')

# Trading cost as $
ew_norebal %>%
  ggplot(aes(x=date, y=commission , fill = ticker)) + 
    theme_set(theme_bw()) + 
    geom_bar(stat = 'identity') +
    ggtitle('3 ETF USD Risk Premia - Commission ($)')

# Trading cost as % of total exposure in instrument
ew_norebal %>%
  mutate(commissionpct = commission / exposure) %>%
  ggplot(aes(x=date, y=commissionpct , fill = ticker)) + 
    geom_bar(stat = 'identity', position = position_dodge()) + 
    theme_set(theme_bw()) + 
    ggtitle('3 ETF USD Risk Premia - Commission as pct of exposure')

# Vanity summary performance metrics
norebal_portfolioreturn <- ew_norebal %>%
  group_by(date) %>%
  summarise(totalequity = sum(exposure),
            totalcommission = sum(commission)) %>%
  ungroup() %>%
  arrange(date) %>%
  mutate(returns = totalequity / lag(totalequity) - 1) 

# Summary performance
norebal_portfolioreturn %>%
  tq_performance(Ra = returns, performance_fun = table.AnnualizedReturns)

# Drawdowns
norebal_portfolioreturn %>%
  tq_performance(Ra = returns, performance_fun = table.DownsideRisk) %>%
  select(MaximumDrawdown)

# Table of trades, exposures etc.
ew_norebal

# Total commssion
totalcommission <- sum(norebal_portfolioreturn$totalcommission)
totalcommission

# Turnover
turnover <- 0
turnover

# Total dollar profit
totalprofit <- tail(norebal_portfolioreturn, 1)$totalequity - initEq

# Trading costs as a percentage of total profit 
costprofit  <- sum(norebal_portfolioreturn$totalcommission) / totalprofit
costprofit

# TODO Rolling performance, portfolio and assets
   
```

# Tab 2: Equal Dollar Weight with Rebalancing

Inputs on tab.

```{r}
initEq <- 10000 # Allow $1000 to 1 million - $100 will be jaws of death...
perShareComm <- 0.005 # Allow 0.001 to 0.02 
minCommPerOrder <- 1 # Allow 0 to 5
rebalFrequency <- 1 # rebalance frequency in months. Allow 1 to 12
capFrequency <- 1 # frequence to capitalise profits. Allow 0 to 12. 0 is "don't capitalise"
```

```{r}
stopifnot(rebalFrequency > 0)

# Create wide data frames for simple share based backtest
wideprices <- monthlyprices %>%
  pivot_wider(date, names_from = 'ticker', values_from = 'close')

rowlist <- list()

cash <- initEq
sharepos <- c(0,0,0)
sharevalue <- c(0,0,0)
equity <- initEq
capEquity <- initEq # Sticky equity amount that we're using to allow change cap frequency

# Iterate through prices and backtest
for (i in 1:(nrow(wideprices))) {
   currentdate <- wideprices[i,1] %>% pull() %>% as.Date()
   currentprice <- as.numeric(wideprices[i,2:4])
   equity <- sum(sharepos * currentprice) + cash
   
   # Update capEquity if it's re-capitalisation rebalance time
   if(capFrequency > 0) {
     if(i %% capFrequency == 0) capEquity <- equity 
   }
   
   # Update position sizing if its position rebalance time 
   if (i == 1 | i %% rebalFrequency == 0) {
    targetshares <- trunc((capEquity / 3) / currentprice) 
   }
   
   trades <- targetshares - sharepos
   tradevalue <- trades * currentprice
   commissions <- abs(trades) * perShareComm
   commissions[commissions < minCommPerOrder] <- minCommPerOrder
   
   # Adjust cash by value of trades
   cash <- cash - sum(tradevalue) - sum(commissions)
   sharepos <- targetshares
   sharevalue <- sharepos * currentprice
   equity <- sum(sharevalue) + cash
   
   # Create data frame and add to list
   row_df <- data.frame(ticker = c('cash', colnames(wideprices[2:4])),
                        date = rep(currentdate, 4),
                        close = c(0,currentprice),
                        shares = c(0,sharepos),
                        exposure = c(cash, sharevalue),
                        sharetrades = c(0, trades),
                        tradevalue = c(-sum(tradevalue), tradevalue),
                        commission = c(0, commissions))
   
   rowlist[[i]] <- row_df
   
}

# Combine list into dataframe
ew_rebal <- bind_rows(rowlist)


# Stacked area chart
ew_rebal %>% 
  ggplot(aes(x=date, y=exposure, fill=ticker)) +
    geom_area() + 
    theme_set(theme_bw()) + 
    ggtitle('3 ETF USD Risk Premia - Equal Weight, Rebalancing')

# Trading cost as $
ew_rebal %>%
  ggplot(aes(x=date, y=commission , fill = ticker)) + 
    theme_set(theme_bw()) + 
    geom_bar(stat = 'identity') +
    ggtitle('3 ETF USD Risk Premia - Commission ($)')

# Trades chart
ew_rebal %>%
  filter(!ticker == 'cash') %>%
  mutate(tradepct = tradevalue / exposure) %>%
  ggplot(aes(x=date, y=tradepct, fill=ticker)) +
    geom_bar(stat = 'identity', position = position_dodge()) + 
    theme_set(theme_bw()) + 
    ggtitle('3 ETF USD Risk Premia - Trades (as % of position size)')

# Vanity summary performance metrics
rebal_portfolioreturn <- ew_rebal %>%
  group_by(date) %>%
  summarise(totalequity = sum(exposure),
            totalcommission = sum(commission)) %>%
  ungroup() %>%
  arrange(date) %>%
  mutate(returns = totalequity / lag(totalequity) - 1) 

# Summary performance
rebal_portfolioreturn %>%
  tq_performance(Ra = returns, performance_fun = table.AnnualizedReturns)

# Drawdowns
rebal_portfolioreturn %>%
  tq_performance(Ra = returns, performance_fun = table.DownsideRisk) %>%
  select(MaximumDrawdown)

# Table of trades, exposures etc.
ew_rebal

# Total commssion
totalcommission <- sum(rebal_portfolioreturn$totalcommission)
totalcommission

# AnnualTurnover
# Calculate as total sell trades divided by mean equity * number of years
totalselltrades <- ew_rebal %>%
  filter(ticker != 'cash',
         tradevalue < 0) %>%
  summarise(sellvalue = sum(tradevalue)) %>%
  pull()

meanequity <- mean(rebal_portfolioreturn$totalequity)

turnover <- -totalselltrades / (meanequity * (year(endDate) - year(startDate)))
turnover

# Total dollar profit
totalprofit <- tail(rebal_portfolioreturn, 1)$totalequity - initEq

# Trading costs as a percentage of total profit 
costprofit  <- sum(rebal_portfolioreturn$totalcommission) / totalprofit
costprofit

# TODO Rolling performance for portfolio and assets
```

# Tab 3 - Simple Risk Parity 

Inputs for tab

```{r}
initEq <- 10000 # Allow $1000 to 1 million - $100 will be jaws of death...
perShareComm <- 0.005 # Allow 0.001 to 0.02 
minCommPerOrder <- 1 # Allow 0 to 5
rebalFrequency <- 1 # rebalance frequency in months. Allow 1 to 12
capFrequency <- 1 # frequence to capitalise profits. Allow 0 to 12. 0 is "don't capitalise"
assetVolTarget <- 0.05 # Allow 0.01 to 0.1
volLookback <- 60 # Allow 5 to 250
```

Calculate vol sizing

```{r}
# Calculate vol target sizing on daily data
theosize_daily <- prices %>%
  group_by(ticker) %>%
  arrange(date) %>%
  mutate(returns = (closeadjusted / dplyr::lag(closeadjusted)) - 1) %>%
  mutate(vol = slider::slide_dbl(.x = returns, .f = sd, .before = volLookback, .complete = TRUE) * sqrt(252)) %>%
  mutate(theosize = lag(assetVolTarget / vol))

# Now enforce a maximum leverage of 1... 
totalsize <- theosize_daily %>%
  group_by(date) %>%
  summarise(totalsize = sum(theosize)) %>%
  mutate(adjfactor = case_when(totalsize > 1 ~ 1/totalsize, TRUE ~ 1))

theosize_constrained <- theosize_daily %>%
  inner_join(totalsize, by = 'date') %>%
  mutate(theosize_constrained = theosize * adjfactor) %>%
  select(ticker, date, closeadjusted, returns, vol, theosize, theosize_constrained) %>%
  na.omit()

# Get the snapshots at the month end boundaries
volsizeprices <- monthlyprices %>%
  inner_join(select(theosize_constrained, ticker, date, theosize_constrained), by = c('ticker','date'))

# Plot theoretical constrained sizing as a function of time
volsizeprices %>%
  ggplot(aes(x=date, y=theosize_constrained, fill=ticker)) +
    geom_area() + 
    theme_set(theme_bw()) + 
    ggtitle('3 ETF USD Risk Premia - Theoretical Constrained Sizing (% of Portfolio Equity')

```

Backtest of vol targeting strategy

```{r}
stopifnot(rebalFrequency > 0)

# Create wide data frames for simple share based backtest
wideprices <- volsizeprices %>%
  pivot_wider(date, names_from = 'ticker', values_from = 'close')

widetheosize <- volsizeprices %>%
  pivot_wider(date, names_from = 'ticker', values_from = 'theosize_constrained')

rowlist <- list()

cash <- initEq
sharepos <- c(0,0,0)
sharevalue <- c(0,0,0)
theosizepct <- c(0,0,0)
targetshares <- c(0,0,0)
equity <- initEq
capEquity <- initEq # Sticky equity amount that we're using to allow change cap frequency

# Iterate through prices and backtest
for (i in 1:(nrow(wideprices))) {
   currentdate <- wideprices[i,1] %>% pull() %>% as.Date()
   currentprice <- as.numeric(wideprices[i,2:4])
   currenttheosize <- as.numeric(widetheosize[i, 2:4])
   equity <- sum(sharepos * currentprice) + cash
   
   # Update capEquity if it's re-capitalisation rebalance time
   if(capFrequency > 0) {
     if(i %% capFrequency == 0) capEquity <- equity 
   }
   
   # Update position sizing if its position rebalance time 
   if (i == 1 | i %% rebalFrequency == 0) {
      targetshares <- trunc((capEquity * currenttheosize) / currentprice)
   }
   
   trades <- targetshares - sharepos
   tradevalue <- trades * currentprice
   commissions <- abs(trades) * perShareComm
   commissions[commissions < minCommPerOrder] <- minCommPerOrder
   
   # Adjust cash by value of trades
   cash <- cash - sum(tradevalue) - sum(commissions)
   sharepos <- targetshares
   sharevalue <- sharepos * currentprice
   equity <- sum(sharevalue) + cash
   
   # Create data frame and add to list
   row_df <- data.frame(ticker = c('cash', colnames(wideprices[2:4])),
                        date = rep(currentdate, 4),
                        close = c(0,currentprice),
                        shares = c(0,sharepos),
                        exposure = c(cash, sharevalue),
                        sharetrades = c(0, trades),
                        tradevalue = c(-sum(tradevalue), tradevalue),
                        commission = c(0, commissions))
   
   rowlist[[i]] <- row_df
   
}

# Combine list into dataframe
volsize_rebal <- bind_rows(rowlist)


# Stacked area chart
volsize_rebal %>% 
  ggplot(aes(x=date, y=exposure, fill=ticker)) +
    geom_area() + 
    theme_set(theme_bw()) + 
    ggtitle('3 ETF USD Risk Premia - Simple Risk Parity')

# Trading cost as $
volsize_rebal %>%
  ggplot(aes(x=date, y=commission , fill = ticker)) + 
    theme_set(theme_bw()) + 
    geom_bar(stat = 'identity') +
    ggtitle('3 ETF USD Risk Premia - Commission ($)')

# Trades chart
volsize_rebal %>%
  filter(!ticker == 'cash') %>%
  mutate(tradepct = tradevalue / exposure) %>%
  ggplot(aes(x=date, y=tradepct, fill=ticker)) +
    geom_bar(stat = 'identity', position = position_dodge()) + 
    theme_set(theme_bw()) + 
    ggtitle('3 ETF USD Risk Premia - Trades (as % of position size)')

# Vanity summary performance metrics
volsize_portfolioreturn <- volsize_rebal %>%
  group_by(date) %>%
  summarise(totalequity = sum(exposure),
            totalcommission = sum(commission)) %>%
  ungroup() %>%
  arrange(date) %>%
  mutate(returns = totalequity / lag(totalequity) - 1) 

# Summary performance
volsize_portfolioreturn %>%
  tq_performance(Ra = returns, performance_fun = table.AnnualizedReturns)

# Drawdowns
volsize_portfolioreturn %>%
  tq_performance(Ra = returns, performance_fun = table.DownsideRisk) %>%
  select(MaximumDrawdown)

# Table of trades, exposures etc.
volsize_rebal

# Total commssion
totalcommission <- sum(norebal_portfolioreturn$totalcommission)
totalcommission

# AnnualTurnover
# Calculate as total sell trades divided by mean equity * number of years
totalselltrades <- volsize_rebal %>%
  filter(ticker != 'cash',
         tradevalue < 0) %>%
  summarise(sellvalue = sum(tradevalue)) %>%
  pull()

meanequity <- mean(volsize_portfolioreturn$totalequity)

turnover <- -totalselltrades / (meanequity * (year(endDate) - year(startDate)))
turnover

# Total dollar profit
totalprofit <- tail(volsize_portfolioreturn, 1)$totalequity - initEq

# Trading costs as a percentage of total profit 
costprofit  <- sum(volsize_portfolioreturn$totalcommission) / totalprofit
costprofit


# TODO rolling performance
```

