---
title: "Risk Premia Strategy Notebook"
output: html_notebook
---

Pull data from bigquery

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, bigrquery, slider, here)

# prices <- query_exec("SELECT *  FROM `rw-algotrader.master_assetclass.assetclass_price`;", 
#                      project = 'rw-algotrader', 
#                      use_legacy_sql = FALSE)

load(here::here("data", "assetclass_prices.RData"))
```

TODO: Don't need this when everything works ok... 

```{r}
indexes <- query_exec("SELECT *  FROM `rw-algotrader.load_assetclass.assetclass_indexes`;", 
                      project = 'rw-algotrader', 
                      use_legacy_sql = FALSE)
```

Dividend adjust these prices... 

TODO - I think adjustedclose on prices now is correct. Validate this.


Plot unadjusted...

```{r}
prices %>% 
  ggplot(aes(x=date, y = close, color = ticker)) + geom_line()
```

Calculate cumulative returns

```{r}
totalreturns <- prices %>%
  arrange(date) %>%
  group_by(ticker) %>%
  mutate(totalreturns = ((close + dividends) / dplyr::lag(close)) - 1) %>%
  select(-closeadjusted) %>%
  na.omit() %>%
  mutate(cumreturns = cumprod(1+totalreturns))

adjratios <- totalreturns %>%
  group_by(ticker) %>%
  summarise(date = max(date)) %>%
  inner_join(totalreturns, by = c('ticker', 'date')) %>%
  mutate(ratio = close / cumreturns) %>%
  select(ticker, ratio)
  
returns <- totalreturns %>%
  inner_join(adjratios, by = 'ticker') %>%
  mutate(adjclose = cumreturns * ratio)
```

```{r}
returns %>%
  ggplot(aes(x=date, y = cumreturns, color = ticker)) + geom_line()
```

```{r}
p <- prices %>% 
  arrange(date) %>%
  group_by(ticker) %>%
  mutate(totalreturns = ((closeadjusted) / dplyr::lag(closeadjusted)) - 1) %>%
  na.omit() %>%
  mutate(cumreturns = cumprod(1+totalreturns)) 

p %>% 
  ggplot(aes(x=date, y = cumreturns, color = ticker)) + geom_line()
  
returns %>% 
  filter(ticker == "GLD") %>% 
  select(ticker, date, totalreturns) %>% 
  inner_join(p, by = c("ticker", "date")) %>% 
  mutate(diff = totalreturns.x - totalreturns.y) %>% 
  ggplot(aes(x = date, y = diff)) + geom_line()
```


Check that returns calculated from the adjusted prices match total returns

```{r}
returns %>%
  mutate(checkreturns = adjclose / lag(adjclose) - 1) %>%
  mutate(check = totalreturns - checkreturns) %>%
  na.omit() %>% 
  ggplot(aes(x = date, y = check)) + geom_line() + facet_wrap(~ticker)
```

# Scatterplots




# Equal weighted backtest 

Calculate monthly returns

```{r}
etfprices <- returns %>% 
  filter(ticker %in% c('VTI','GLD','TLT')) 
```

```{r}
library(tidyquant)

# Calculate monthly returns
monthly_ew <- etfprices %>%
  tq_transmute(select = adjclose,
               mutate_fun = periodReturn,
               period = 'monthly',
               col_rename = 'return') %>%
  group_by(date) %>%
  summarise(portfolio = sum(1/3 * return))
  
monthly_ew

monthly_ew %>%
  ggplot(aes(x=date, y=cumprod(1+portfolio))) +
  geom_line()


monthly_ew %>%
  tq_performance(Ra = portfolio, performance_fun = table.AnnualizedReturns)

monthly_ew %>%
  tq_performance(Ra = portfolio, performance_fun = table.AnnualizedReturns)
```

# Daily equal weight

```{r}
daily_ew <- etfprices %>%
  group_by(date) %>%
  summarise(portfolio = mean(totalreturns))

daily_ew %>%
  ggplot(aes(x=date, y = cumprod(1 + portfolio))) + geom_line()

daily_ew %>%
  tq_performance(Ra = portfolio, performance_fun = table.AnnualizedReturns)

daily_ew %>% filter(date >= '2010-01-01') %>%
  tq_performance(Ra = portfolio, performance_fun = table.AnnualizedReturns)
```


# Vol targeting on the individual assets

Daily rebalance

```{r}
library(slider)
vol_target <- 0.05 # size each to 5% annualised vol

volsize <- etfprices %>%
  mutate(vol20 = slider::slide_dbl(.x = totalreturns, .f = sd, .before = 60, .complete = TRUE) * sqrt(252)) %>%
  mutate(vol20size = lag(vol_target / vol20)) %>%
  mutate(vol20returns = vol20size * totalreturns) %>%
  na.omit()

volsizeportfolio <- volsize %>%
  group_by(date) %>%
  summarise(portfolio = sum(vol20returns))

volsizeportfolio %>%
  ggplot(aes(x=date, y = cumprod(1 + portfolio))) + geom_line()

volsizeportfolio %>%
  tq_performance(Ra = portfolio, performance_fun = table.AnnualizedReturns)

volsizeportfolio %>% filter(date >= '2010-01-01') %>%
  tq_performance(Ra = portfolio, performance_fun = table.AnnualizedReturns)
```

# Monthly rebalance

```{r}
monthreturns <- etfprices %>%
  tq_transmute(select = adjclose,
               mutate_fun = periodReturn,
               period = 'monthly',
               col_rename = 'return') %>%
  inner_join(volsize, by = c('ticker', 'date')) %>%
  select(ticker, date, vol20size, return)

adjustmentfactors <- monthreturns %>%
  group_by(date) %>%
  summarise(totalweight = sum(vol20size)) %>%
  mutate(adjfactor = case_when(totalweight > 1 ~ 1/totalweight, TRUE ~ 1))

monthstrategyreturns <- monthreturns %>%
  inner_join(adjustmentfactors, by = c('date')) %>%
  mutate(constrainedsize = vol20size * adjfactor) %>%
  mutate(weightedreturn = constrainedsize * return)

month_rebal <- monthstrategyreturns %>%
  group_by(date) %>%
  summarise(portfolio = sum(weightedreturn))

month_rebal %>%
  filter(date >= '2010-01-01') %>%
  ggplot(aes(x=date, y = cumprod(1 + portfolio))) + geom_line()

month_rebal %>%
  tq_performance(Ra = portfolio, performance_fun = table.AnnualizedReturns)

month_rebal %>% filter(date >= '2010-01-01') %>%
  tq_performance(Ra = portfolio, performance_fun = table.AnnualizedReturns)

monthstrategyreturns %>%
  group_by(date) %>%
  summarise(totalweight = sum(constrainedsize)) %>%
  ggplot(aes(x=date, y = totalweight)) + geom_line()
  
```

# Get the portfolio returns

```{r}
library(lubridate)
dump <- month_rebal %>%
  filter(date >= '2010-01-01') %>%
  mutate(date = floor_date(date, 'month')) 
 # select(date, returns = portfolio) %>%
 #  tq_performance(Ra = portfolio, performance_fun = table.AnnualizedReturns)


saveRDS(dump, file = 'rp-returns-7vol.RDS')
```

# Scatterplot of returns 

```{r}
monthreturns %>%
  mutate(pastreturns = lag(return)) %>%
  ggplot(aes(x=pastreturns, y = return)) +
  geom_point() + 
  facet_wrap(~ticker) + 
  ggtitle('Monthly Returns vs Past Month Returns')
```

# Scatterplot of volatility

```{r}
monthreturns %>%
  mutate(vol20  = vol_target / vol20size) %>%
  mutate(pastvol20 = lag(vol20)) %>%
  ggplot(aes(x=pastvol20, y = vol20)) +
  geom_point() + 
  facet_wrap(~ticker) + 
  ggtitle('Volatility vs Past Month Volatility')

```

